---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First we need to unzip and load data into the project:
```{r echo = TRUE}
activity <- read.csv(unzip("activity.zip"))
```

Let`s look on the first ten rows and information about each colomn:
```{r echo = TRUE}
head(activity, 10)
summary(activity)
str(activity)
```

As we see, the second colomn has Factor type. Let`s make it Date:
```{r echo = TRUE}
activity$date <- as.Date(activity$date)
str(activity)
```

Now looks better!

## What is mean total number of steps taken per day?

First we will make a table with dates and number of steps per day
```{r echo = TRUE}
library(plyr)
library(dplyr)
steps_per_day <- ddply(activity,.(date),summarize,total_steps=sum(steps, na.rm=TRUE))
head(steps_per_day)
```

There is a histogram of the total number of steps per day:
```{r echo = TRUE}
library(ggplot2)

g <- ggplot(steps_per_day, 
            aes(x=date, 
                y=total_steps))
g <- g + geom_bar(fill="#0000ff", stat="identity")
g <- g + labs(x = "Date") 
g <- g + labs(y = "Total steps") 
g <- g + labs(title="Total number of steps per day")
g
```

The mean and median of total number of steps are:
```{r echo = TRUE}
steps_mean <- mean(steps_per_day$total_steps)
steps_median <- median(steps_per_day$total_steps)
```
mean = ```r steps_mean```;

median = ```r steps_median```;

## What is the average daily activity pattern?

We will make a table of 5-minutes intervals and the average number of steps taken, averaged across all days:
```{r echo = TRUE}
avg_interval <- ddply(activity,.(interval),summarize,avg_steps=mean(steps, na.rm=TRUE))
head(avg_interval)
```


Then make a plot:
```{r echo = TRUE}

g <- ggplot(avg_interval, 
            aes(x=interval, 
                y=avg_steps))
g <- g + geom_line()
g <- g + labs(x = "Interval") 
g <- g + labs(y = "Average number of steps") 
g <- g + labs(title="average daily activity")
g
```

Then find a 5-minute interval with maximum number of steps:
```{r echo = TRUE}
max_interval <- avg_interval$interval[which.max(avg_interval$avg_steps)]
```

So, the ```r max_interval``` 5-minute interval contains the maximum number of steps.

## Imputing missing values

Let`s calculate a number of missing values in our dataset:
```{r echo = TRUE}
NA_number <- sum(is.na(activity$steps))
```
There are ```r NA_number``` missing values.

Now we will fill all NA values by the mean for that 5-minute interval

```{r echo = TRUE}
missing <- which(is.na(activity))
new_activity <- activity

new_activity$steps[missing] <- round(avg_interval$avg_steps[activity$interval[missing] / 5 + 1])
head(new_activity, 10)
new_steps_per_day <- ddply(new_activity,.(date),summarize,total_steps=sum(steps, na.rm=TRUE))
head(new_steps_per_day)
```

There is a histogram of the total number of steps per day without NA values:
```{r echo = TRUE}
g <- ggplot(new_steps_per_day, 
            aes(x=date, 
                y=total_steps))
g <- g + geom_bar(fill="#0000ff", stat="identity")
g <- g + labs(x = "Date") 
g <- g + labs(y = "Total steps") 
g <- g + labs(title="Total number of steps per day")
g
```

New mean and median of total number of steps are:
```{r echo = TRUE}
new_steps_mean <- mean(new_steps_per_day$total_steps)
new_steps_median <- median(new_steps_per_day$total_steps)
```
new_mean = ```r new_steps_mean```;

new_median = ```r new_steps_median```;

There is a difference between that values with and without NA:

mean - new_mean = ```r steps_mean - new_steps_mean```;

median - new_median = ```r steps_median - new_steps_median```;

## Are there differences in activity patterns between weekdays and weekends?

Next step we will create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day:
```{r echo = TRUE}
Sys.setlocale("LC_TIME", "English_Australia.1252")
new_activity$w <- factor(ifelse(weekdays(new_activity$date,abbreviate=TRUE) %in% c("Sat","Sun"), "weekend","weekday"))
head(new_activity, 10)
```

A panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis):
```{r echo = TRUE}
library(reshape2)
activity_melt <- melt(new_activity, id.vars = c("w", "interval"), measure="steps")
steps_mean <-dcast(activity_melt, w + interval ~ variable, fun = mean)

library(lattice)
xyplot(steps ~ interval | w, data = steps_mean, type='l', layout = c(1, 2), ylab="number of steps")
```